---
title: "Empirical and expected homorepeat counts in Swiss-Prot"
date: "January 14, 2016"
output: html_document
---


```{r include=FALSE}
rm(list = ls(all = TRUE))
gc()
setwd("/home/matteo/polybox/MSc_ACLS/swissrepeat/results")
source("helpers.R")
```


```{r, echo=FALSE, eval=TRUE}
aa_ignore = c("B", "X", "Z", "O", "U")
homo_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/empirical_and_expected/swissprot.csv"
homo_all = load_homorepeat_data(homo_path, aa_ignore, "all")

homo_disorder_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/empirical_and_expected/mobidb_regions_iupl.csv"
homo_d_all = load_homorepeat_data(homo_disorder_path, aa_ignore , "disorder")

homo_order_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/empirical_and_expected/mobidb_regions_iupl_inverted.csv"
homo_o_all = load_homorepeat_data(homo_order_path, aa_ignore, "order")

homo = rbind(homo_all, homo_d_all, homo_o_all)

homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/swissprot.csv"
homo_exp_all = load_expected_homorepeat_frequencies(homo_exp_path, aa_ignore, "all")

homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/mobidb_consensus.csv"
homo_exp_d = load_expected_homorepeat_frequencies(homo_exp_path, aa_ignore, "disorder")

homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/mobidb_consensus_inverse.csv"
homo_exp_o = load_expected_homorepeat_frequencies(homo_exp_path, aa_ignore, "order")

homo_exp = rbind(homo_exp_all, homo_exp_d, homo_exp_o)

# Needed to add \n to .csv files for R to read them properly
# aa_counts_eukaryota_disorder_path = "results/swiss_prot_aa_frequencies/swissprot_disorder_kingdomwise/Eukaryota.csv"
aa_counts_eukaryota_disorder_path = "/results/amino_acid_frequencies/Eukaryota.csv"
aa_counts_eukaryota_disorder = load_amino_acid_counts(aa_counts_eukaryota_disorder_path, "disordered", "Eukaryota", aa_ignore)
# aa_counts_eukaryota_order_path = "results/swiss_prot_aa_frequencies/swissprot_disorder_inverse_kingdomwise/Eukaryota.csv"
aa_counts_eukaryota_order_path = "/results/amino_acid_frequencies/Eukaryota.csv"
aa_counts_eukaryota_order = load_amino_acid_counts(aa_counts_eukaryota_order_path, "ordered", "Eukaryota", aa_ignore)
# aa_counts_bacteria_disorder_path = "results/swiss_prot_aa_frequencies/swissprot_disorder_kingdomwise/Bacteria.csv"
aa_counts_bacteria_disorder_path = "/results/amino_acid_frequencies/Bacteria.csv"
aa_counts_bacteria_disorder = load_amino_acid_counts(aa_counts_bacteria_disorder_path, "disordered", "Bacteria", aa_ignore)
# aa_counts_bacteria_order_path = "results/swiss_prot_aa_frequencies/swissprot_disorder_inverse_kingdomwise/Bacteria.csv"
aa_counts_bacteria_order_path = "/results/amino_acid_frequencies/Bacteria.csv"
aa_counts_bacteria_order = load_amino_acid_counts(aa_counts_bacteria_order_path, "ordered", "Bacteria", aa_ignore)

aa_counts_eukaryota = compare_amino_acid_counts(aa_counts_eukaryota_order, aa_counts_eukaryota_disorder)
aa_counts_bacteria = compare_amino_acid_counts(aa_counts_bacteria_order, aa_counts_bacteria_disorder)

aa_counts = rbind(aa_counts_eukaryota, aa_counts_bacteria)
```

Sidenote: "Sometimes, the specific identity of an amino acid cannot be determined unambiguously. Certain protein sequencing techniques do not distinguish among certain pairs. Thus, these codes are used:

    Asx (B) is "asparagine or aspartic acid"
    Glx (Z) is "glutamic acid or glutamine"
    Xle (J) is "leucine or isoleucine"

In addition, the symbol X is used to indicate an amino acid that is completely unidentified." [From Wiki](https://en.wikipedia.org/wiki/Proteinogenic_amino_acid)


### Show amino acid frequencies in ordered and disordered regions
TODO: Where is the NA value coming from?
```{r}
aa_counts[which(is.na(aa_counts$aa)),]
aa_counts <- aa_counts[which(!is.na(aa_counts$aa)),]

```
Don't know where they're from, but since there only very few, let's remove them.

```{r, echo=FALSE, eval=TRUE}
p = ggplot(aa_counts, aes(x=frequency, y=aa, colour=Superkingdom, size=total_frequency)) + 
  geom_point()+
  facet_wrap(~ type, scales = "free")
beautifier(p)

aa_counts_tmp = subset(aa_counts, type=="ordered")
p = ggplot(aa_counts_tmp, aes(x=log10relative_frequency_ordered_divided_by_disordered, y=aa, colour=Superkingdom, size=total_frequency)) + geom_point()
p = p + facet_wrap(~ type, scales = "free")
beautifier(p)
```
The frequency of each AA (ordered by their decreasing disorder-promoting potential (top: ordered, bottom: disordered)) from ordered and disordered regions of Bacteria and Eukaryota.
From Uversky2013: 
In fact, in comparison with ordered pro-
teins, IDPs/IDPRs are characterized by noticeable biases in their 
amino acid compositions, containing less of so-called 
“order-promoting” residues (cysteine, tryptophan, isoleucine, 
tyrosine, phenylalanine, leucine, histidine, valine, asparagines 
and methionine, which are mostly hydrophobic residues which 
are commonly found within the hydrophobic cores of foldable 
proteins) and more of “disorder-promoting” residues (lysine, 
glutamine, serine, glutamic acid and proline, which are mostly 
polar and charged residues, which are typically located at the 
surface of foldable proteins)

### Show average amino acid frequencies in ordered and disordered regions per kingdom:
```{r, echo=FALSE, eval=TRUE}
aa_frequency_summary = ddply(aa_counts, .(Superkingdom, type), summarize, std_frequency = sd(frequency))
aa_frequency_summary
```


### Calculate the expected ratio of homorepeats in ordered and disordered regions, given their empirical amino acid distributions. Compare this theoretical result to the above data.

```{r, echo=FALSE, eval=TRUE}
## Test whether empirical and expected numbers are sensible. 

# "Release 2015_11 of 11-Nov-15 of UniProtKB/Swiss-Prot contains 549832 sequence entries,comprising 196078138 amino acids abstracted from 240597 references. "
# ftp://ftp.uniprot.org/pub/databases/uniprot/previous_releases/release-2015_11/knowledgebase/UniProtKB_SwissProt-relstat.html


# The expected and empirical values need to add up to approximately the same value:

# For the entire of Swiss-Prot, this value is the total number of amino acids in Swiss-Prot, 196078138:
# Counting the seq length in Swiss-Prot (Retrieved with BioPython): 196219159 - currently, cannot explain difference to official numbers.
# In Swiss-Prot annotations for aa counts: 196219157
sum(homo_all[homo_all$type=="expected","repeat_region_length"]) 
sum(homo_all[homo_all$type=="empirical","repeat_region_length"])

# At current, we ignore ~60 sequence, were disorder annotations (MobiDB) are only available for outdated Swiss-Prot sequences. Therefore, character counts don't add up to the total Swiss-Prot length.
sum(homo_d_all[homo_d_all$type=="expected","repeat_region_length"]) 
sum(homo_d_all[homo_d_all$type=="empirical","repeat_region_length"]) 

sum(homo_o_all[homo_o_all$type=="expected","repeat_region_length"]) 
sum(homo_o_all[homo_o_all$type=="empirical","repeat_region_length"]) 

## Do empirical and expected numbers add up to the same value?
# -> Not yet, however to a low margin. Needs checking.
## Do numbers for order and disorder add up to the total of Swiss-Prot?
# -> Not yet, as we ignore a couple of dozen sequence lacking current disorder annotations
```


#### Absolute homorepeat counts
```{r, echo=FALSE, eval=TRUE}
homo_q = subset(homo_all, aa=="Q")
p = ggplot(homo_q, aes(x=n, y=log10_count_rounded, colour=type)) + geom_point()
p = p + facet_wrap(~ Kingdom, scales = "free")
p = p + ggtitle('Amino acid: Q')
beautifier(p)

homo_q = subset(homo_all, aa %in% c("N", "Q", "S", "E") & type == "empirical")
p = ggplot(homo_q, aes(x=n, y=log10_count_rounded, colour=aa)) + geom_point()
p = p + facet_wrap(~ Kingdom, scales = "free")
p = p + ggtitle('Empirical data')
beautifier(p)


show_homorepeat_counts <- function(data, kingdom, nmin, nmax){
  homo_sub = subset(data, Kingdom==kingdom & n >= nmin & n <= nmax)
  p = ggplot(homo_sub, aes(x=n, y=log10_count_rounded, colour=type)) + geom_point(size=1.5)
  p = p + facet_wrap(~ aa)
  beautifier(p)
}

show_homorepeat_counts(homo_all, "Eukaryota", 1, 50)
show_homorepeat_counts(homo_all, "Bacteria", 1, 50)
show_homorepeat_counts(homo_all, "Viruses", 1, 50)
show_homorepeat_counts(homo_all, "Archaea", 1, 30)
```

#### Empirical homorepeat counts normalized by expected homorepeat counts.
```{r, echo=FALSE, eval=TRUE}
# This solution ONLY works, if homo is correctly ordered. Otherwise, find a smarter split/divide/merge method!
homo_expected = subset(homo, type == "expected")
homo_empirical = subset(homo, type == "empirical")

homo_empirical$relative_count = homo_empirical$count / homo_expected$count
homo_empirical$log10relative_count = log10(homo_empirical$relative_count)
homo_empirical = subset(homo_empirical)

show_empirical_homorepeat_counts_relative_to_expected <- function(data, kingdom, nmin, nmax){
  homo_empirical = subset(data, Kingdom==kingdom & n >= nmin & n <= nmax & set != "all")
  homo_empirical = homo_empirical[complete.cases(homo_empirical),]
  p = ggplot(homo_empirical, aes(x=n, y=log10relative_count, colour=set)) + geom_point()
  p = p + facet_wrap(~ aa)
  p = p + scale_x_continuous(breaks = (seq(nmin, nmax, by = round((nmax/max(nmin,1))/5))))
  p = p + geom_hline(yintercept=0)
  beautifier(p)
}

show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Eukaryota", 0, 50)
show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Bacteria", 0, 50)
show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Viruses", 0, 50)
show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Archaea", 0, 50)

show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Eukaryota", 0, 10)
show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Bacteria", 2, 6)
show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Viruses", 2, 6)
show_empirical_homorepeat_counts_relative_to_expected(homo_empirical, "Archaea", 2, 10)
```

#### Empirical homorepeat frequencies (= counts / region lengths)
```{r, echo=FALSE, eval=TRUE}
show_empirical_homorepeat_counts_relative_to_total_seq_length <- function(data, kingdom, nmin, nmax){
  homo_empirical = subset(data, Kingdom==kingdom & n >= nmin & n <= nmax & set != "all" & type == "empirical")
  p = ggplot(homo_empirical, aes(x=n, y=frequency, colour=set)) + geom_point()
  p = p + facet_wrap(~ aa)
  #p = p + scale_y_continuous(limits = c(0, 10))
  p = p + scale_x_continuous(breaks = (seq(nmin, nmax, by = round((nmax/max(nmin,1))/5))))
  p = p + geom_hline(yintercept=0)
  beautifier(p)
}

show_empirical_homorepeat_counts_relative_to_total_seq_length(homo, "Eukaryota", 0, 50)
show_empirical_homorepeat_counts_relative_to_total_seq_length(homo, "Bacteria", 0, 30)
show_empirical_homorepeat_counts_relative_to_total_seq_length(homo, "Viruses", 0, 30)
show_empirical_homorepeat_counts_relative_to_total_seq_length(homo, "Archaea", 0, 30)

show_empirical_homorepeat_counts_relative_to_total_seq_length(homo, "Eukaryota", 0, 10)
```



#### Fraction of homorepeats in IDRs of all homorepeats (normalized and not normalized by aa frequencies)
```{r, echo=FALSE, eval=TRUE}
# homo$frequency is the homorepeat count normalised by sequence length.
# We need to calculate (homo$frequency [Disorder] / ( homo$frequency [Order] + homo$frequency [Disorder]))

homo_d_frequency = homo[homo$type=="empirical" & homo$set=="disorder", c("Kingdom", "aa", "n", "frequency")]
colnames(homo_d_frequency)[4] <- "d_frequency"
homo_o_frequency = homo[homo$type=="empirical" & homo$set=="order", c("Kingdom", "aa", "n", "frequency")]
colnames(homo_o_frequency)[4] <- "o_frequency"

homo_frequency = merge(homo_d_frequency, homo_o_frequency, by = c("Kingdom","aa", "n"))
homo_frequency$relative_fraction_disordered_homorepeats = homo_frequency$d_frequency / (homo_frequency$o_frequency + homo_frequency$d_frequency)
homo_frequency$type = "empirical"
homo_frequency = homo_frequency[, c("Kingdom", "aa", "n", "relative_fraction_disordered_homorepeats", "type")]

tmp = subset(homo_frequency, Kingdom=="Eukaryota" & n < 15)

# homo_expected$frequency is the expected frequency of homorepeats in the data.
# We need to calculate (homo_expected$frequency [Disorder] / ( homo_expected$frequency [Order] + homo_expected$frequency [Disorder]))

homo_d_frequency = homo[homo$type=="expected" & homo$set=="disorder", c("Kingdom", "aa", "n", "frequency")]
colnames(homo_d_frequency)[4] <- "d_frequency"
homo_o_frequency = homo[homo$type=="expected" & homo$set=="order", c("Kingdom", "aa", "n", "frequency")]
colnames(homo_o_frequency)[4] <- "o_frequency"

homo_frequency_exp = merge(homo_d_frequency, homo_o_frequency, by = c("Kingdom","aa", "n"))
homo_frequency_exp$relative_fraction_disordered_homorepeats = homo_frequency_exp$d_frequency / (homo_frequency_exp$o_frequency + homo_frequency_exp$d_frequency)
homo_frequency_exp$type = "expected_(bernoulli)"
homo_frequency_exp = homo_frequency_exp[, c("Kingdom", "aa", "n", "relative_fraction_disordered_homorepeats", "type")]


# Merge homo_normalized_sequence_length and homo_normalized_exp,  
homo_normalized = rbind(homo_frequency_exp, homo_frequency)

# Visualise data

show_disorderd_vs_ordered_homorepeat_counts <- function(data, kingdom, nmin, nmax){
  homo_empirical = subset(data, Kingdom==kingdom & n >= nmin & n <= nmax)
  p = ggplot(homo_empirical, aes(x=n, y=relative_fraction_disordered_homorepeats, colour=type)) + geom_point(alpha=0.8)
  p = p + facet_wrap(~ aa)
  #p = p + scale_y_continuous(limits = c(0, 10))
  p = p + scale_x_continuous(breaks = (seq(nmin, nmax, by = round((nmax/max(nmin,1))/5))))
  p = p + geom_hline(yintercept=0)
  beautifier(p)
}

show_disorderd_vs_ordered_homorepeat_counts(homo_normalized, "Eukaryota", 0, 15)
show_disorderd_vs_ordered_homorepeat_counts(homo_normalized, "Archaea", 0, 15)
show_disorderd_vs_ordered_homorepeat_counts(homo_normalized, "Bacteria", 0, 15)
show_disorderd_vs_ordered_homorepeat_counts(homo_normalized, "Viruses", 0, 15)

```




#### Expected homorepeat counts in an unbound sequence given the amino acid frequencies
```{r, echo=FALSE, eval=TRUE}
# Add up the frequencies for all amino acids
homo_exp_sum = ddply(homo_exp, .(kingdom, n, set), summarize, expected_homorepeat_frequency_sum_over_all_aa = sum(expected_frequency))
homo_exp_sum$log10_expected_homorepeat_frequency_sum_over_all_aa = log10(homo_exp_sum$expected_homorepeat_frequency_sum_over_all_aa)

homo_exp_sum_tmp = subset(homo_exp_sum, n <= 25)
p = ggplot(homo_exp_sum_tmp, aes(x=n, y=log10_expected_homorepeat_frequency_sum_over_all_aa, colour=set)) + geom_point() + facet_wrap(~ kingdom) + scale_y_continuous(breaks = (seq(-26, 0, by = 2)))
beautifier(p)


homo_exp_sum = subset(homo_exp_sum, n <= 6)
p = ggplot(homo_exp_sum, aes(x=n, y=log10_expected_homorepeat_frequency_sum_over_all_aa, colour=set)) + geom_point() + facet_wrap(~ kingdom) + scale_y_continuous(breaks = (seq(-26, 0, by = 2)))
beautifier(p)
```

