---
title: 'TODO: Homorepeats in Swissprots'
output:
  html_document: default
  pdf_document: default
date: "March, 2019"
---

# Housekeeping
```{r Housekeeping, echo=FALSE, message=FALSE, warning=FALSE}
# Before executing this file, do: setwd("path/to/your/git/on/swissrepeat/results")
setwd("/home/matteo/polybox/MSc_ACLS/swissrepeat/results")
source("local_config.R")
setwd(paste0(local_base_path,"/results"))

rm(list = ls(all = TRUE))
gc()
source("helpers.R")

# colour setup:
#library(RColorBrewer); display.brewer.all() # to display available colour palettes
colour_count = 13 # alternative: length(unique(sp_gathered$Kingdom))
getPalette = colorRampPalette(brewer.pal(9, "Dark2"))
```

# Data Loading 
```{r Data Loading , echo=FALSE, eval=TRUE}
tr_path = paste0("results", local_path_separator, "tr_annotations", local_path_separator, "tr_annotations.csv")
tr_all = load_tr_annotations(tr_path)

sp_path = paste0("data", local_path_separator, "swissprot_annotations.tsv")
sp_all = load_swissprot(sp_path, tr_all)

# Add meta_data from sp_all to tr_all. -> Do a left join.
tr_all_sp = merge(x = tr_all, y = sp_all, by = "ID", all.x = TRUE)

# Add disorder information from modiDB
discoor_path = paste0("results", local_path_separator, "disorder_annotations", local_path_separator, "mobidb_coordinates.csv")
discoor_all = load_disorder_annotations(discoor_path)

discoor_all$disorder_region_length =  (discoor_all$end - discoor_all$start) + 1
discoor_all$center = floor(discoor_all$start + (discoor_all$disorder_region_length/2))
discoor_all_sp = merge(x = discoor_all, y = sp_all, by = "ID", all.x = TRUE)

# Remove eventual regions with incorrect coordinates
discoor_all_sp = discoor_all_sp[(discoor_all_sp$center<discoor_all_sp$Length),]

#Aggregate disordered regions by protein, calculate disorder residues count in a protein
sp_protein= ddply(discoor_all, .(ID), summarize,
                  disorder_count=(sum(disorder_region_length)))
sp_protein = merge(x = sp_protein, y = sp_all, by = "ID", all.x = TRUE)


aa_ignore = c("B", "X", "Z", "O", "U", "*","-",".","+", "other")
homo_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/empirical_and_expected/swissprot.csv"
homo_all = load_homorepeat_data(homo_path, aa_ignore, "all")

homo_disorder_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/empirical_and_expected/mobidb_consensus.csv"
homo_d_all = load_homorepeat_data(homo_disorder_path, aa_ignore, "disorder")

homo_order_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/empirical_and_expected/mobidb_consensus_inverse.csv"
homo_o_all = load_homorepeat_data(homo_order_path, aa_ignore, "order")

homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/swissprot.csv"
# homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/swissprot_kingdomwise.csv"
# couldn't find the file "swissport_kingdomwise.csv" and don't know where it's generated... however swissprot exists in this directory and seems to work.
homo_exp_all = load_expected_homorepeat_frequencies(homo_exp_path, aa_ignore, "all")

homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/mobidb_consensus.csv"
homo_exp_d = load_expected_homorepeat_frequencies(homo_exp_path, aa_ignore, "disorder")

homo_exp_path = "results/empirical_and_expected_homorepeat_counts/swiss_prot_homorepeat_frequencies/expected_on_unbound_sequence/mobidb_consensus_inverse.csv"
homo_exp_o = load_expected_homorepeat_frequencies(homo_exp_path, aa_ignore, "order")
```
# Homorepeats
```{r}
nrow(tr_all_sp) # no. of all TRs in swissprot
nrow(homo_all) # no. of all homorepeats (CHECK! -> wrong! See here: length(unique(tr_all_sp$ID[which(tr_all_sp$has_homo_tr == TRUE) ])) )

nrow(homo_all)/nrow(tr_all_sp)

nrow(homo_all[which(homo_all$Kingdom == "Eukaryota"),])/nrow(tr_all_sp)

nrow(homo_all[which(homo_all$Kingdom == "Eukaryota"),])/nrow(tr_all_sp[which(tr_all_sp$Superkingdom == "Eukaryota"),])

nrow(homo_all[which(homo_all$Kingdom == "Eukaryota"),])/nrow(homo_all)

str(sp_all)
```


### Fig. 5a
```{r, echo=FALSE, eval=TRUE, fig.width=10, fig.height=5}
# for all AAs
homo_q = subset(homo_all, type == "empirical")
p = ggplot(homo_q, aes(x=n, y=log10_count_rounded, colour=aa)) +
  geom_point(size=2)+
  facet_wrap(~ Kingdom, scales = "fixed")+
  ggtitle('Empirical data')+
  labs(x = "Repeat unit number",
       y = expression(log[10]~count~(rounded)))+
  scale_color_manual(values = morecolors2)+
  guides(colour=guide_legend(override.aes = list(size = 3)),
         size=FALSE)
p = beautifier(p, x.axis.text.angle = 0)
p <- paper.figure(p, x.axis.text.angle = 0)
p
if( save) {
  ggsave(paste0(pathImages, "fig5a_allAA", figureFormat), width=12, height=8, dpi = 300)
}



# For a subset of AAs
homo_q = subset(homo_all, aa %in% c("N", "Q", "S", "E") & type == "empirical")

p = ggplot(homo_q, aes(x=n, y=log10_count_rounded, colour=aa)) +
  geom_point(size=2)+
  facet_wrap(~ Kingdom, scales = "fixed")+
  ggtitle('Empirical data')+
  labs(x = "Repeat unit number",
       y = expression(log[10]~count~(rounded)))+
  scale_color_manual(values = cols1,
                     name = "AA")+
  guides(colour=guide_legend(override.aes = list(size = 3)),
         size=FALSE)
p = beautifier(p, x.axis.text.angle = 0)
p <- paper.figure(p, x.axis.text.angle = 0)
p
if( save) {
  ggsave(paste0(pathImages, "fig5a", figureFormat), width=12, height=8, dpi = 300)
}
```
Fig. 5a). Count of homorepeats in Swiss-Prot in four Superkingdoms for different repeat unit number ($n <= 50$, equivalent to repeat length) for hydrophilic Asparagine (N), Glutamine (Q), Serine (S) and Glutamic acid (E). Homorepeats with large $n$ seem to mostly pertain to the Eukaryotes. 

### Fig. 5b
```{r, echo=FALSE, eval=TRUE, fig.width=10, fig.height=6}
# Plot empirical vs expected counts
show_homorepeat_counts <- function(data, kingdom, nmin, nmax, cols1){
  homo_sub = subset(data, Kingdom==kingdom & n >= nmin & n <= nmax)
  p = ggplot(homo_sub, aes(x=n, y=log10_count_rounded, colour=type)) + 
    geom_point(size=1.5)+
    facet_wrap(~ aa)+
    labs(x = "Repeat unit number",
         y = expression(log[10]~count~(rounded)))+
    guides(colour=guide_legend(override.aes = list(size = 3)), 
           size=FALSE)+
    theme(panel.spacing.x=unit(1.5, "lines"),panel.spacing.y=unit(1, "lines"))+
    scale_color_manual(values = c(cols1[1], cols1[4])) # c("#D46A6A", "#D49A6A", "#407F7F", "#55AA55")
  beautifier(p, x.axis.text.angle = 0)
  paper.figure(p, x.axis.text.size = 18, x.axis.text.angle = 45, x.axis.text.hjust = 1)
}

p = show_homorepeat_counts(homo_all, "Eukaryota", 1, 50, cols1)
p
if( save) {
  ggsave(paste0(pathImages, "fig5b", figureFormat), width=12, height=8, dpi = 300)
}

```
Fig. 5b). Empirical and expected count of homorepeats in Swiss-Prot Eukaryotes ($n <= 50$). Amino acids are ordered by their propensity to promote structural order.


### Fig. 5c
```{r, echo=FALSE, eval=TRUE, fig.width=10, fig.height=6}
# Plot empirical counts for order vs disorder.
n_chars_swiss_prot = sum(homo_all[homo_all$type=="empirical","repeat_region_length"])
homo_all$set = "all"
homo_d_all$set = "disorder"
homo_o_all$set= "order"
homo_all$count_relative_sequence_length = homo_all$count
homo_d_all$count_relative_sequence_length = homo_d_all$count/sum(homo_d_all[homo_d_all$type=="empirical","repeat_region_length"]) * n_chars_swiss_prot
homo_o_all$count_relative_sequence_length = homo_o_all$count/sum(homo_o_all[homo_o_all$type=="empirical","repeat_region_length"]) * n_chars_swiss_prot

homo = rbind(homo_all, homo_d_all, homo_o_all)
homo$set = as.factor(homo$set)
homo$log10_count_relative_sequence_length = log10(homo$count_relative_sequence_length)

show_empirical_homorepeat_counts_relative_to_total_seq_length <- function(data, kingdom, nmin, nmax, cols1){
  homo_empirical = subset(data, Kingdom==kingdom & n >= nmin & n <= nmax & set != "all" & type == "empirical")
  p = ggplot(homo_empirical, aes(x=n, y=log10_count_relative_sequence_length, colour=set)) + 
    geom_point(size=1.5) +
    facet_wrap(~ aa)+
    # scale_y_continuous(limits = c(0, 10))+
    scale_x_continuous(breaks = (seq(nmin, nmax, by = round((nmax/max(nmin,1))/5))))+
    geom_hline(yintercept=0)+
    labs(x = "Repeat unit number",
         y = expression(log[10]~count~(rounded)))+
    guides(colour=guide_legend(override.aes = list(size = 3)), 
           size=FALSE)+
    theme(panel.spacing.x=unit(1.5, "lines"),panel.spacing.y=unit(1, "lines"))+
    scale_color_manual(values = c(cols1[1], cols1[4])) # c("#D46A6A", "#D49A6A", "#407F7F", "#55AA55")
  beautifier(p, x.axis.text.angle = 0)
  paper.figure(p, x.axis.text.size = 18, x.axis.text.angle = 45, x.axis.text.hjust = 1)
}

p = show_empirical_homorepeat_counts_relative_to_total_seq_length(homo, "Eukaryota", 0, 50, cols1)
if( save) {
  ggsave(paste0(pathImages, "fig5c", figureFormat), width=12, height=8, dpi = 300)
}
p
```
Fig. 5c). Empirical count of homorepeats in Swiss-Prot Eukaryotes ($n <= 50$) for ordered and disordered regions (consensus MobiDB annotations, no minimum length cut-off.). Amino acids are ordered by their propensity to promote structural order.

